---
title: Sec X. Measures of salmon habitat quality
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load the package
library("knitcitations")

## clean out any existing .bib files
cleanbib()

## set pandoc for formatting
options("citation_format" = "pandoc")
```



### task

either some references for a habitat quality index or 0.5-1 page of text describing the index and highlighting any potential work in constructing the index that a Sea Grant Fellow can be tasked with.


# Cites


`r citet("10.1111/fwb.13358")`

However, nonlinear relationships may exist between the amount of available habitat and the productivity of stream dwelling salmonids `r citep("10.1111/1365-2656.12845")`

`r citet("10.1111/1365-2656.12845")`


`r citet("10.1139/cjfas-2016-0406")`


We will also explore the possibility of including the interactive effects of location on the landscape and projected effects of climate change. For example, Battin et al. `r citet("10.1073/pnas.0701685104")` found that future climate change and coordinated habitat restoration working in synergy are likely to cause a shift in the distribution of salmon in Puget Sound. 

Importance of temperature for growth opportunities. habitat quality warmer areas 
`r citet("10.1002/ecs2.3266")`

habitat connectivity index (HCI) based upon upstream passability and segment length. `r citet("10.1080/20442041.2018.1544817")`


Pess `r citep("10.1007/s11160-013-9339-1")`

```{r}
## create ref cited section
write.bibtex(file = "references.bib")
```

