\documentclass[12pt]{elsarticle}
\usepackage{newtxtext}
\usepackage[margin=1in]{geometry}
%\pagestyle{empty}
\usepackage{natbib}
\bibliographystyle{unsrt}
\usepackage{color}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{titlesec}
\usepackage{colortbl}
\titleformat{\section}[runin]{\normalfont\bfseries}{\thesection.}{3pt}{}


\begin{document}

\begin{center} \textbf{DATA SHARING PLAN} \end{center}

%All data primary data produced and secondary data utilized in this project will be made available to the general public after appropriate quality control checks and a minimal delay (not more than 1 year after the project end date). The data will be made available through a public GitHub repository. In addition to data sharing, the GitHub repository will be used to make public replication materials for any published articles resulting from this project.
\begin{enumerate}

\item	Title of the Proposal: Maximizing returns on investing in barrier culvert removal in Washington state\\

\item	Name of the lead PI: Sunny L. Jardine\\

\item		Contact Information\\

Sunny L. Jardine\\ University of Washington\\ Marine Studies Building \# 229\\ Email: jardine@uw.edu\\ Phone: (206) 616-4952\\

\item		Dataset Description(s) (required answer)
What data will the dataset(s) contain? This includes descriptive details on data types, inclusion of metadata, data format(s), collection times / date ranges, etc. What name(s), if any, will be designated to the dataset(s)?

\begin{itemize}
\item A geospatial dataset of predicted cost for all known barrier culverts in the Washington State injunction area along with all predictor variables. The data will be stored as a shapefile (.shp) and include metatdata in the form of an accompanying .html or .Rmd file.
\item A geospatial dataset of usual and accustomed fishing areas for each Tribal Nation in the Washington State injunction area. The data will be stored as a shapefile (.shp) and include metatdata in the form of an accompanying .html or .Rmd file.
\item A geospatial dataset of the habitat quantity (linear miles) and quality, for each of 5 Pacific salmon species and steelhead, gained by removing each known barrier culvert assuming no downstream blockages, i.e. the quantity and quality of habitat for each species of interest between a barrier and the next upstream barrier. The data will be stored as a shapefile (.shp) and include metatdata in the form of an accompanying .html or .Rmd file.
\item A catalog of prioritization methods, and variables included in the prioritization indices, for all entities (e.g. counties) in the Washington State injunction area. The data and associated metadata will be stored in an Excel spreadheet (.xlsx).\\
\end{itemize}

\item		Do you agree to release all data no later than 2 years after the end-date of the project?\\
Yes\\

\item		Issues (required answer)
Are there any legal, access, retention, etc. issues anticipated for the dataset? If yes, please explain.\\
No\\

\item		Data Size
What will be the estimated size of the dataset? Please report estimated number of MB, GB, TB, etc., collected.\\
750 MB\\

\item		 Data Format
What format will the dataset utilize? (i.e., Excel file, model code, audio/video recording, etc.)\\
Shapefiles and Excel files (see answer to question X above).\\


\item		Ownership (required answer)
Who will own the dataset, if not the lead PI?\\
NA\\

\item		Post-Processing
What post-processing, QA/QC will this dataset undergo? Who will be responsible for performing this post-processing and QA/QC to prepare the dataset for its deposition into a repository?\\
The cost and habitat datasets will be reviewed by PI Jardine. The dataset on Tribal fishing areas and catalog of prioritization methods will be reviewed by co-PI VanDeynze. In each review, the code generating the data will be reviewed (where applicable) and a random sample of observations will be selected for validation.\\

\item		Preservation Plan (required answer) 
What data repositories will be used to host the dataset? If none, how will the data be preserved?\\
The data will be hosted on a publicly-available GitHub repository.\\
 
\item		Products
 Will any information or data products be developed from this dataset? How will the related costs be supported? Which organization(s) will be producing these products?\\
 A decision support tool will be developed from this dataset. Washington Sea Grant funds will be used to develop the tool. \\

\item		Other Comments
Are there any additional comments related to the data that will results from your Sea Grant-funded study?\\
No\\


\end{enumerate}

\end{document}