---
title: "Culvert Cost Models with Spatially Explicit Data"
author: "B. Van Deynze"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
   html_document:
      number_sections: true
      toc: true
      toc_float:
         collapsed: true
---


```{r include=F}
# Prepare environment and data ----
rm(list = ls())

library(equatiomatic)
library(MASS)
library(tidyverse)
library(janitor)
library(here)
library(scales)
library(knitr)
library(kableExtra)
library(broom)
library(margins)
library(ggeffects)
library(forcats)
library(readxl)
library(mctest)
library(ggcorrplot)


opts_chunk$set(echo=FALSE)

# Wraps long subtitles in ggplot
wrapper <- function(label, dev_width = dev.size("in")[1], dev_scaler = 12)  {   
  paste(strwrap(label, dev_width * dev_scaler), collapse = "\n") 
}

# Introduction and theory ----
```
```{r }
```

# Incorporating costs in conservation planning  
**Niche:** use of cost data in conservation plans  

- Just as benefits, variability in costs can be large  
  - Including spatial variability and variability across scope and scale of project  
  - Understanding this variability can improve planning outcomes  

***Key cites***  
**Babcock et al. 1997:** https://doi.org/10.2307/3147171  
  
- Describes relative efficiency of management rules under different joint distributions of costs and benefits  
- Alternative targeting instruments considered incl. cost-targeting, benefit-targeting, and ratio-targeting (cost per benefit targeting)  
- Relative variability of benefits and costs, and correlation between the two, determine effects of sub-optimal targeting  
  

```{r fig.width=10, fig.height=10, echo=F, message=F, warning=F}
set.seed(123456)
sigma = matrix( 
  c(2, 0, 0, 1), # the data elements 
  nrow=2,              # number of rows 
  ncol=2,              # number of columns 
  byrow = TRUE
)        # fill matrix by rows
xy <- mvrnorm(100, c(0, 0), sigma)
# cov(xy)
# cor(xy)
xy <- data.frame(xy)
   
ggplot() +
  geom_text(
    aes(
      x = c(-3, 1, -3, 1),
      y = c(3, 3, -2.95, -2.95)
    ),
    label =
      c(
        "I. High-benefit, low-cost",
        "II. High-benefit, high-cost",
        "III. Low-benefit, low-cost",
        "IV. Low-benefit, high-cost"
      ),
    hjust = 0,
    size = 5
  ) +
  # geom_rect(
  #   aes(
  #     xmax = c(0.5, 1, 0.5, 1),
  #     xmin = c(0, 0.5, 0, 0.5),
  #     ymax = c(1, 1, 0.5, 0.5),
  #     ymin = c(0.5, 0.5, 0, 0)
  #   ),
  #   fill = c("green", "yellow", "yellow", "red")
  # ) +
  geom_point(
    aes(
      x = X1,
      y = X2
    ),
    data = xy
  ) +
  # stat_ellipse(
  #   aes(
  #     x = X1,
  #     y = X2
  #   )
  # ) +
  geom_abline(
    aes(color = "Ratio cost (C/B) targeting",
    slope = 1, intercept = 0),
    size = 1.2
  ) +
  geom_vline(
    aes(color = "Cost targeting",
    xintercept = 0),
    size = 1.2
  ) +
  geom_hline(
    aes(color = "Benefit targeting",
    yintercept = 0),
    size = 1.2
  ) +
  labs(
    title = "Potential projects in cost-benefit space",
    subtitle = wrapper("Lines illustrate thresholds for which projects would be selected under different targeting schemes; Projects above and to the left of each line would be selected under each scheme"),
    x = "Cost",
    y = "Benefit",
    caption = "Recreation of figures 1-3 of Babcock et al. (1997)"
  ) +
  ggthemes::theme_clean() +
  scale_color_discrete("Targeting scheme") +
  theme(
    plot.title.position = "plot",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.position = "bottom"
  ) +
  coord_fixed(xlim = c(-3, 3), ylim = c(-3, 3))
```

**Naidoo et al. 2006:** https://doi.org/10.1016/j.tree.2006.10.003  
  
- Types of costs: acquisition, management, transaction (and opportunity, damage costs)  
  - (Can be continuous or one-off)  
- Often based on non-monetary proxies  
  - Most often area  
  - Sometimes weighted but in often arbitrary ways  
  - Efficiency gains from incorporating costs  
  
**Gap**
  
- Past looks at culverts have focused on benefits and used simplified cost models  
- Past looks at conservation costs have focused on land acquisition costs rather than restoration efforts  
  - Unique features of culvert improvement in PNW: upstream land access model, lots of streams/roads, large variation in slope and stream size  
- Timely b/c Washington culvert case  
  
**Research approach**
  
- Examine variability in cost levels and drivers of costs across culvert projects in PNW using statistical model  
- Compare levels and variability of costs to (possibly several) benefit measures  
- Apply model to extant culverts to compare costs/benefit distributions over...  
  - Space: where are high benefit, low cost culverts?  
  - Observed projects vs. all culverts: what kind of decision rule is distribution of projects consistent with?  


*RQ1:* How much variability is there in costs for culvert improvements?  
  
- Over space?  
- For observed projects vs. potential projects?  
- Relative to variability in benefits? (And implications for planning rules/future research)  
  
*RQ2:* What are drivers of culvert improvement costs?  
  
- Economic drivers: economies of scale, transaction costs  
- Geophysical drivers: stream features, terrain features  


```{r include=F}

# Load NLCD key
key_nlcd <-
  read_xlsx(
    here(
      "data/Culverts spatial overlays v 06Aug2020.xlsx"
    ), 
    sheet = 3
  ) %>% 
  as_tibble() %>%
  clean_names() %>%
  rename(
    nlcd_current_class = class,
    nlcd_current_fullclass = classification
  ) %>%
  mutate(across(where(is_character), str_to_sentence)) %>%
  filter(across(description, ~!str_detect(., "Alaska only"))) %>%
  select(-description)

# Load data
df_culv <-
  read_csv(here("output", "culverts_pure_modelling.csv")) %>%
  mutate(
    # project_year = ordered(project_year),
    project_source = relevel(factor(project_source), ref = "OWRI"),
    basin = relevel(factor(basin), ref = "SOUTHERN OREGON COASTAL"),
    fips = factor(fips),
    state_fips = factor(state_fips),
    # here_class = ordered(here_class)
  ) %>%
  left_join(
    key_nlcd, 
    by = c("nlcd_current" = "value")
  )
  
  
names(df_culv)

# Data description ----
```
```{r echo=F}
```

# Data description

The unit of observation in our data is a **culvert work site**. These data
include all unique work sites associated with a culvert action in the PNSHP
data between 2001 and 2015. Each work site is associated with a *project*, a
set of geographic coordinates, and the number of culverts at the site.
Projects are associated with a year, a reporting source, and a unique cost.
We also calculate the number of culverts associated with the project (methods
found
[here](https://vandeynze.github.io/salmon_culverts/culverts_summary.html#Appendix:_Creating_a_Consistent_Project-level_Culvert_Count)).
Note that a project may be associated with multiple related work sites,
though `r tabyl(df_culv$n_worksites) %>% slice(1) %>% pull(percent) %>%
percent` of work sites are uniquely identified to a project.

Dependent and explanatory variables included in the empirical model are described below,
including a brief justification for inclusion. A more in-depth exploration of
these variables can be found in [this
report](https://https://vandeynze.github.io/salmon_culverts/spatial_variables_summary.html).  

```{r }
options(knitr.kable.NA = '')
df_culv %>%
  drop_na(emp_agforest, slope, slope_deg) %>%
  select(
    `Cost per culvert ($USD2019)` = cost_per_culvert,
    `Number of culverts (count)` = n_culverts,
    `Mean distance between work sites (m)` = dist_mean,
    `Stream slope (%)` = slope,
    `Bankfull width (m)` = bankfull_width,
    `Paved road` = here_paved,
    `Road class` = here_class,
    `Terrain slope (deg)` = slope_deg,
    `Land cover class` = nlcd_current_class,
    `Housing density (units per sq. km)` = hdens_cat,
    `Construction employment (jobs)` = emp_const,
    `Ag/forestry employment (jobs)` = emp_agforest,
    `Basin` = basin,
    `Year` = project_year,
    `Reporting source` = project_source
  ) %>%
  mutate(
    `Road class` = factor(`Road class`),
    `Paved road` = factor(`Paved road`),
    `Land cover class` = factor(`Land cover class`),
    Year = ordered(Year)
  ) %>%
  summarize(
    across(where(is.numeric), list(Mean = mean, `Std. dev.` = sd), .names = "{col}_{fn}", na.rm = TRUE),
    across(where(is.factor), list(`Number of levels` = nlevels), .names = "{col}_{fn}")
  ) %>%
  pivot_longer(everything(), names_to = c("Variable", "stat"), names_sep = "_") %>%
  rowwise() %>%
  mutate(across(where(is.numeric), ~(format(signif(., digits = 3), scientific = FALSE, big.mark = ",")))) %>%
  pivot_wider(Variable, names_from = stat, values_from = value) %>%
  # print(n = Inf)
  kable(
    caption = paste0("Descriptive statistics (n = ", comma(nrow(df_culv %>% drop_na(emp_agforest, slope, slope_deg))), ")"),
    # escape = FALSE,
  ) %>%
  kable_styling("hover", fixed_thead = TRUE)
```

## Cost estimates  
  
**Cost per culvert ($USD, 2019)** is our primary dependent variable. This
variable can also be interpreted as the *project average costs* at the work
site. This variable is constructed by dividing the provided project costs by
the number of culverts associated with the project.  
  
## Stream hydrological features  

- **Stream slope (% grade)**: slope of stream at road crossing can require more
expensive crossing design; identified via COMID matching with NHDPlus
attributes.  
- **Bankfull width (m)**: bankfull width is the preferred measure of stream width
at road crossing, accounting for potential width during high-water events;
identified via COMID matching with NHDPlus attributes.  

```{r echo=F}
```

## Road features  

- **Road paved (indicator)**: modification of a paved road is more expensive;
may also proxy for higher traffic volumes; measured via HERE road data for
nearest object.  
- **Road class (categorical)**: wider roads with more traffic are expected to
be more expensive; measured via HERE road data for nearest object; classes range from 2 (largest) to 5 (smallest)


```{r echo=F}
key_here <- read_xlsx(here("data/Culverts spatial overlays v 20Aug2020.xlsx"), sheet = 3) %>%
  as_tibble() %>%
  mutate(Classification = str_to_sentence(Classification), Description = str_to_sentence(Description)) %>%
  bind_rows(
    tibble(
      "Classification" = "Functional class",
      "Value" = 6,
      "Description" = '"Roads" associated with work sites > 150m from the nearest HERE road for here_class_badmatch.'
    )
  )

key_here %>%
  filter(Classification == "Functional class") %>%
  arrange(Classification) %>%
  kable(caption = "Value key for HERE road variables") %>%
  kable_styling() %>%
  collapse_rows(
    1,
    valign = "top",
    row_group_label_position = "stacked",
    headers_to_remove = 1
  )
```

## Terrain features  

- **Terrain slope (degrees)**: steeper terrain is expected to require more
expensive projects; measured by the
[GTOPO30](https://www.usgs.gov/centers/eros/science/usgs-eros-archive-digital-elevation-global-30-arc-second-elevation-gtopo30?qt-science_center_objects=0#qt-science_center_objects)
grid cell the work site falls in.

- **Land cover (categorical)**: different land covers may be associated more
expensive projects (e.g. less accessible sites in forest, difficult soils in
welands, etc.); identified via cover with work site coordinates and NLCD land
cover layer for nearest available year; here we use the broader NLCD Group
definition rather than the more detailed classification (see below).  


```{r echo=F}
(
  key_nlcd <-
    read_xlsx(
      here(
        "data/Culverts spatial overlays v 06Aug2020.xlsx"
      ), 
      sheet = 3
    ) %>% 
    as_tibble() %>%
    clean_names("sentence") %>% 
    mutate(across(where(is_character), str_to_sentence)) %>%
    rename(Group = Class) %>%
    filter(across(Description, ~!str_detect(., "Alaska only")))
) %>%
  # print(n = 20)
  knitr::kable() %>%
  kableExtra::kable_styling()
```
```{r echo=F}
```

## Population features  

- **Housing density (units per sq. km)**: more parcels near work site
introduces complexities related to site access and available areas for
staging, etc.; measured for the immediate catchment of stream identified via
matching with NHDPlus attribute data.  
- **Employment in ag/forestry (jobs in county)**: availability of skilled
labor may reduce project costs; employment data from
county the work site is located in via [County Business
Patterns](https://www.census.gov/programs-surveys/cbp.html) data.  
- **Employment in construction (jobs in county)**: see above.  
  

```{r echo=F}
```

## Scale and scope controls  

- **Number of culverts associated with project (count):** addressing multiple
culverts under the same project may provide scale benefits, but might also
increase complexity; measured via PNSHP database.  
- **Distance between project work sites (m)**: more dispersed work sites
under a single project may increase project costs due to increased
transportation costs (and time); measured as the log of the average euclidean distance
between work sites for multiple work site projects (plus one to define the variable at zero); a dummy for single work
site projects is also included to disentangle scale and distance effects.  
- **Action type (categorical):** PNSHP distinguishes between culvert removals
and culvert installations, in addition to culvert improvements (the dominate
category); we expect removals to be cheapest, followed by improvements and
installations; dummies are included when a project includes one or more
culverts flagged as either removals or installations.  
  

```{r echo=F}
```

## Fixed effects  

- **Year**: the year the project was completed  
- **Basin**: the basin (HUC6) where the work site is located  
- **Reporting source**: the reporting source for the project  

```{r }
# Simple correlations across variables ----
# +
```

# Simple correlations across variables  

Here we present a couple measures of correlation between potential continuous
explanatory variables. The figure below show provides Pearson's correlation
coefficients for each pair of continuous explanatory variables included in
the initial models, along side the dependent variable. Also included is the
Variance Inflation Factor for each variable as calculated when all presented
variables are included in a simple log-linear model, with cost per culvert as
the dependent variable. Because of previously mentioned high correlation
between housing density and population density, we include only housing
density.

```{r fig.width=10, fig.height=10, echo=F, message=F, warning=F}
# Grab only variables needed for corr. plot
df_corr <-
  df_culv %>%
  select(
    # employment vars
    `Employment, construction` = emp_const, `Employment, ag/forestry` = emp_agforest,
    # pop vars
    `Housing density` = hdens_cat,
    # popdens_cat,
    # stream vars
    `Stream slope` = slope, `Bankfull width` = bankfull_width,
    `Terrain slope` = slope_deg, 
    `Mean distance between sites` = dist_mean, `Number of culverts` = n_culverts,
    `Cost per culvert` = cost_per_culvert
  ) %>%
  drop_na()

mat_corr <-
  df_corr %>%
  cor() %>%
  as.matrix()


mat_vifs <- imcdiag(lm(log(`Cost per culvert`) ~ ., df_corr), method = "VIF")[[1]] %>% as.matrix
mat_vifs <- rbind(mat_vifs, c(NA, NA))
diag(mat_corr) <- mat_vifs[,1] %>% round(1)
mat_corr[upper.tri(mat_corr)] <- NA

ggcorrplot(
  mat_corr,
  lab = TRUE, lab_size = 3,
  colors = c("#6D9EC1", "white", "#E46726")
) + 
  theme(
    plot.title.position = "plot",
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(color = "#ADADAD", linetype = "dashed"),
    axis.text.x = element_text(angle = 60, vjust = 1),
    # plot.margin = unit(c(0.5, 0, 1.5, 0), "cm")
  ) +
  scale_fill_gradient2(
    name = "Corr. coef. (r)",
    limits = c(-1, 1),
    low = "#F8766D",
    mid = "#FFFFFF",
    high = "#619CFF",
    na.value = "#EAEAEA"
  ) +
  ggtitle(
    "Correlations among potential covariates",
    "VIF along diagonal"
  ) +
  coord_fixed(
    # ylim = c(0, 20),
    # xlim = c(0, 20),
    expand = FALSE,
    clip = "off"
  )

```
```{r }
```

It looks like mean distance between work sites and the number of culverts is
positively correlated, and both are negatively correlated with average
project costs. We would expect distance to increase costs but the
number of culverts to decrease costs (due to economies of scale), all else
equal. The strong positive correlation between number of culverts and
distance between sites. Disentangling these effects should be possible with
multiple regression.

Stream slope is negatively correlated with bankfull width, which means wider
streams tend to be less steep. Stream slope is also positively correlated
with terrain slope, as mentioned earlier. None of the three are strongly
correlated with costs.

Finally, the two employment variables are positively correlated, and
construction employment is positively correlated with housing density.
Ag/forestry employment is weakly positively correlated with measures that
indicate more rugged terrain such as stream and terrain slope. Housing density
and to a lesser degree construction employment are positively correlated with
costs.

No variables have particularly large VIFs, suggesting little potential for
error-inflating multicollinearity. (A high VIF indicates that the variable is
strongly correlated with the other variables in the model, leading to
inflated standard errors and limiting the model's usefulness for prediction
or inference.)

```{r estimate, include=F}
# Estimate models ----
# Full model
mod_full <- 
  lm(
    log(cost_per_culvert) ~
      # Scale/scope of project controls: number of culverts, distance between work sites, type of culvert work
      n_culverts + log(dist_mean+1) + factor(I(n_worksites == 1)) +
      action_fishpass_culvrem_prj + action_fishpass_culvinst_prj +
      # Stream features at work site: slope, bankfull width
      slope * bankfull_width + 
      # Road features at work site: paved, road class
      here_paved + factor(here_class) +
      # Physical features of work site: terrain slope, land cover
      slope_deg + factor(nlcd_current_class) +
      # Population features: housing density, jobs in construction, jobs in ag/forestry
      hdens_cat + emp_const + emp_agforest +
      # Fixed effects
      basin + factor(project_year) + project_source,
    df_culv
  )

# Remove some variables
# Remove scale/scope controls
# mod_drop_scale <-
#   mod_full %>% update(. ~ . - (n_culverts + log(dist_mean+1) + factor(I(n_worksites == 1))))
# mod_drop_scope <-
#   mod_full %>% update(. ~ . - (n_culverts + log(dist_mean+1) + factor(I(n_worksites == 1)) + action_fishpass_culvrem_prj + action_fishpass_culvinst_prj))
# 
# # Remove physical work site features
# mod_drop_phys <-
#   mod_full %>% update(. ~ . - (slope_deg + factor(nlcd_current)))
# 
# # Remove pop features
# mod_drop_pop <-
#   mod_full %>% update(. ~ . - (hdens_cat + emp_const + emp_agforest))
# mod_drop_emp <-
#   mod_full %>% update(. ~ . - (emp_const + emp_agforest))

# Remove everything but streams and roads
# mod_drop_streamsonly <-
#   mod_full %>% update(
#     . ~ . - (
#       n_culverts + log(dist_mean + 1) + factor(I(n_worksites == 1)) +
#         action_fishpass_culvrem_prj + action_fishpass_culvinst_prj +
#         here_paved + factor(here_class) +
#         slope_deg + factor(nlcd_current) +
#         hdens_cat + emp_const + emp_agforest
#     )
#   )
# 
# mod_drop_roadsonly <-
#   mod_full %>% update(
#     . ~ . - (
#       n_culverts + log(dist_mean + 1) + factor(I(n_worksites == 1)) +
#         action_fishpass_culvrem_prj + action_fishpass_culvinst_prj +
#         slope * bankfull_width +
#         slope_deg + factor(nlcd_current) +
#         hdens_cat + emp_const + emp_agforest
#     )
#   )
# 
# mod_drop_roadsandstreams <-
#   mod_full %>% update(
#     . ~ . - (
#       n_culverts + log(dist_mean + 1) + factor(I(n_worksites == 1)) +
#         action_fishpass_culvrem_prj + action_fishpass_culvinst_prj +
#         slope_deg + factor(nlcd_current) +
#         hdens_cat + emp_const + emp_agforest
#     )
#   )

# Remove fixed effects
mod_nofe <-
  mod_full %>% update(. ~ . - (basin + factor(project_year) + project_source))
mod_nofe_onlysource <-
  mod_full %>% update(. ~ . - (basin + factor(project_year)))
mod_nofe_onlyyear <-
  mod_full %>% update(. ~ . - (basin + project_source))
mod_nofe_onlybasin <-
  mod_full %>% update(. ~ . - (factor(project_year) + project_source))
mod_nofe_nobasin <-
  mod_full %>% update(. ~ . - basin)
mod_nofe_noyear <-
  mod_full %>% update(. ~ . - factor(project_year))
mod_nofe_nosource <-
  mod_full %>% update(. ~ . - project_source)

# Focus on certain basins
mod_basins_wore <-
  mod_full %>% update(data = df_culv %>% filter(basin %in% c("SOUTHERN OREGON COASTAL", "NORTHERN OREGON COASTAL", "WILLAMETTE")))
mod_basins_wwash <-
  mod_full %>% update(data = df_culv %>% filter(basin %in% c("WASHINGTON COASTAL", "PUGET SOUND")))
mod_basins_core <-
  mod_full %>% update(data = df_culv %>% filter(basin %in% c("WASHINGTON COASTAL", "PUGET SOUND", "SOUTHERN OREGON COASTAL", "NORTHERN OREGON COASTAL", "WILLAMETTE")))

# Focus on certain sources
mod_sources_owri <-
  mod_full %>% update(. ~ . - project_source, data = df_culv %>% filter(project_source == "OWRI"))
mod_sources_warco <-
  mod_full %>% update(. ~ . - project_source, data = df_culv %>% filter(project_source == "WA RCO"))
mod_sources_core <-
  mod_full %>% update(. ~ . - project_source, data = df_culv %>% filter(project_source %in% c("WA RCO", "OWRI", "HABITAT WORK SCHEDULE", "BLM", "REO")))

```
```{r methods, echo=F, message=F, warning=F}
```

# Estimation  

We estimate log-linear models estimated via OLS, with the average project
cost as the dependent variable and the work site as the unit of observation.
Stream slope and bankfull width are interacted in this specification.
Recommendations in culvert engineering reports indicate that more expensive
culvert designs are particularly necessary when both of these variables are
extreme, and an interaction term can capture this effect.


```{r }
# Giving up on rendering the equation for now
# extract_eq(mod_full)
```


In addition to the fully specified model (mod_full), we present thirteen
alternative models that include different fixed effects configurations or
only sub-samples of the data focused on basin or reporting source criteria.
For basins, we provide results estimated on a "core" group representing the
five most frequently represented basins (Washington Coastal, Puget Sound,
Southern Oregon Coastal, Northern Oregon Coastal, Willamette), as well as
that core separated into basins primarily in Western Washington and Oregon
respectively. For reporting sources, we focus on a similar "core" group (
Washington Recreation and Conservation Office, Oregon Water Resources
Inventory, Habitat Work Schedule, Bureau of Land Management, Regional
Ecosystem Office [an interagency group]), as well as versions estimated only
on OWRI and WA RCO projects.

The resulting coefficients for the fixed effects and categorical variables,
when exponentiated, can be interpreted as the ratio of average costs for that
group relative those of the base group. Results for continuous variables are
presented as exponentiated average marginal effect of a single standard
deviation change, which can be interpreted as the ratio of costs relative to
a work site with a standard deviation lower for the variable.  

## Coefficient estimates  

```{r echo=F}
# Present model estimates ----
mods <- mget(ls(patter = "mod_"))
mods <- mods[c(4, 1:3, 5:14)]
mods_pars <-
  map_df(mods, tidy, .id = "model") %>%
  complete(model, term) %>%
  mutate(
    stars = case_when(
      p.value < 0.01 ~ "***",
      p.value < 0.05 ~ "**",
      p.value < 0.1 ~ "*",
      TRUE ~ ""
    ),
    full.est = if_else(
      is.na(estimate),
      "&#8210;",
      paste0(
        format(signif(estimate, 3), scientific = FALSE, drop0trailing = TRUE, trim = TRUE),
        stars,
        "<br>(", format(signif(std.error, 3), scientific = FALSE, drop0trailing = TRUE, trim = TRUE), ")"
      )
    ),
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "action_fishpass_culvinst_prj" ~ "Culvert installation (dummy)",
      term == "action_fishpass_culvrem_prj" ~ "Culvert removal (dummy)",
      term == "n_culverts" ~ "Number of culverts",
      term == "dist_mean" ~ "Mean distance between work sites",
      term == "log(dist_mean + 1)" ~ "Mean distance between work sites, log",
      term == "factor(I(n_worksites == 1))TRUE" ~ "Single work site (dummy)",
      term == "slope" ~ "Stream slope",
      term == "bankfull_width" ~ "Bankfull width",
      term == "here_pavedY" ~ "Road paved (dummy)",
      term == "slope_deg" ~ "Terrain slope",
      term == "hdens_cat" ~ "Housing density",
      term == "emp_const" ~ "Construction employment",
      term == "emp_agforest" ~ "Ag/forestry employment",
      term == "slope:bankfull_width" ~ "Stream slope X bankfull width",
      TRUE ~ term
    ),
    term = str_replace(term, "project_source", "Project source: "),
    term = str_replace(term, "basin", "Basin: "),
    term = str_replace(term, "factor[(]project_year[)]", "Year: "),
    term = str_replace(term, "factor[(]nlcd_current_class[)]", "Land cover: "),
    term = str_replace(term, "factor[(]here_class[)]", "Road class: ")
  ) %>%
  select(
    model, term, full.est
  ) %>%
  pivot_wider(id_cols = model, names_from = term, values_from = full.est) %>%
  select(
    model,
    Intercept,
    # Stream features
    "Stream slope",
    "Bankfull width",
    "Stream slope X bankfull width",
    # Road features
    starts_with("Road"),
    # Land features
    "Terrain slope",
    starts_with("Land cover"),
    # Pop features
    "Housing density",
    ends_with("employment"),
    # Scale/scope features
    "Number of culverts",
    starts_with("Mean distance between "),
    "Single work site (dummy)",
    starts_with("Culvert "),
    starts_with("Project source: "),
    starts_with("Year: "),
    starts_with("Basin: ")
  )
mods_stats <-
  map_df(mods, glance, .id = "model") %>%
  mutate(
    `Adj. R2` = signif(adj.r.squared, 3),
    AIC = round(AIC, 1),
    BIC = round(BIC, 1),
    N = df + df.residual
  ) %>%
  select(model, `Adj. R2`, AIC, BIC, N) %>%
  mutate_all(~as.character(.))
mods_pars %>%
  left_join(mods_stats, by = "model") %>%
  pivot_longer(-model, names_to = "Term", values_to = "Value") %>%
  pivot_wider(id_cols = Term, names_from = model, values_from = Value) %>%
  select(Term, mod_full, contains("nofe"), contains("basins"), contains("sources")) %>%
  # kable() %>%
  kable(
    caption = "Pilot models",
    escape = FALSE,
    align = 
      paste0(
        "l", 
        paste(
          rep(
            "c",
            length(mods)
            ), 
          collapse = ""
          )
        )
    ) %>%
  kable_styling("hover", fixed_thead = TRUE) %>%
  column_spec(1 , bold = TRUE) %>%
  add_header_above(
    c(
      " " = 1,
      " " = 1,
      "Alternative fixed effects" = 7,
      "Sub-sample: basins" = 3,
      "Sub-sample: sources" = 3
    )
  ) %>%
  row_spec(c(ncol(mods_pars) - 1, ncol(mods_pars) + ncol(mods_stats) - 2), extra_css = "border-bottom: 1px solid") %>%
  add_footnote("* p < 0.1, ** p < 0.05, *** p < 0.01", notation = "none") %>%
  scroll_box(height = "800px")
```

## Model fit discussion  

The full model has an adjusted R-squared of 0.44, indicating a decent model
fit. The version of the model with no fixed effects has an adjusted R-squared
of 0.276, indicating that a significant amount of variability is explained by
the additional explanatory variables. When fixed effect categories are
removed, we can check with fixed effects explain the most variation relative
to each other. It looks like reporting source accounts for the most
variation, followed by basin then year.  

When the model is fit only on culverts in the "core" basins, adjusted
R-squared improves slightly, indicating the model performs better in these
basins relative to the ones with less representation in the sample. This is
particularly true for the Western Oregon basins where the model still has a
pretty high R-squared. It's harder to say whether the reduced R-squared for
the Western Washington basins is due to weaker model fit or simply small
sample size.  

When the model is fit only on the "core" reporting sources, R-squared drops
significantly. This might be evidence that including the additional sources
is important for improving the overall fit of the model, even though less
than 100 observations are lost.  

```{r echo=F, message=F, warning=F}
# Plot figures ----
```

# Model visualizations  

## Slope and bankfull width interaction effect  


```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}
# ____ Slope and width effect space ----


predict_cost_interaction <-
  function(model, var1 = "slope", var2 = "bankfull_width", lims1 = c(0, 0.3), lims2 = c(0, 50), by1 = 0.01, by2 = 5) {
    model %>%
      ggpredict(
        c(
          paste0(var1, " [", paste(lims1, collapse = ":"), " by=", by1, "]"),
          paste0(var2, " [", paste(lims2, collapse = ":"), " by=", by2, "]")
        )
      ) %>%
      tibble %>%
      mutate(
        !!var2 := as.numeric(as.character(group)),
        !!var1 := x,
        x = NULL,
        group = NULL
      )
  }

map_df(mods["mod_full"], predict_cost_interaction, .id = "model") %>%
  # filter(model == "mod_full") %>%
  ggplot() +
  aes(
    x = slope,
    y = bankfull_width,
    z = predicted
  ) +
  geom_contour_filled(
    breaks = c(
      0,
      5000,
      7500,
      10000,
      12500,
      15000,
      20000,
      40000,
      80000,
      160000,
      Inf
    )
  ) +
  scale_fill_brewer(
    name = wrapper("Predicted cost per culvert ($USD)"),
    direction = -1,
    palette = "Spectral",
    labels = c(
      "$5,001 to $7,500",
      "$7,501 to $10,000",
      "$10,001 to $12,500",
      "$12,501 to $15,000",
      "$15,001 to $20,000",
      "$20,001 to $40,000",
      "$40,001 to $80,000",
      "$80,001 to $160,000",
      "Over $160,000"
    )
  ) +
  # facet_wrap("model", nrow = round(sqrt(length(mods)))) +
  geom_point(
    aes(
      x = slope,
      y = bankfull_width,
      z = NULL
    ),
    data = df_culv,
    color = "grey30",
    alpha = 0.3,
    size = 0.2
  ) +
  theme(
    # legend.position = "bottom" 
  ) +
  coord_fixed(0.3/50) +
  labs(
    title = "Predicted average costs by bankfull width (m) and slope (% grade)",
    subtitle = wrapper("Predictions based on full model with other continuous variables at means and categorical variables at their modes; points represent underlying observations"),
    x = "Slope",
    y = "Bankfull width"
  )

df_culv %>%
  ggplot() +
  aes(
    x = slope,
    y = bankfull_width,
    color = cost_per_culvert
  ) +
  geom_point() +
  scale_color_fermenter(
    name = "Cost per culvert ($USD)",
    breaks = c(
      0,
      5000,
      7500,
      10000,
      12500,
      15000,
      20000,
      40000,
      80000,
      160000,
      Inf
    ),
    direction = -1,
    labels = label_dollar(),
    palette = "Spectral",
    # guide = "legend"
  ) +
  theme(
    legend.position = c(0.9, 0.8),
    legend.key.size = unit(1, "cm"),
    plot.title.position = "plot"
  ) +
  labs(
    title = "Observed average cost by bankfull width (m) and slope (% grade)",
    subtitle = wrapper("Difficult to see pattern in raw data, likely due large underlying variation resulting from fixed effects (different basins, years, etc.)"),
    x = "Slope",
    y = "Bankfull width"
  )
```
```{r }
```

Work sites are distributed along a slope - bankfull width convex curve, with
few projects both high slope and high width. This pattern mirrors the cost
contours over slope - bankfull width space. Comparing the observed projects
to other culverts in the Washington or Oregon inventories will show whether
this relationship exists for all culverts or whether projects were selected
along the cost curve. That is, would projects that did not occur exist in the
upper-right space?  

### Continuous variables  

To display how costs vary across the continuous explanatory variables, we
plot average marginal effects. These are scaled to a standard deviation
change for each variable. Average marginal effects are calculated with other
variables held at their means, which is most relevant for slope and bankfull
width for which the model allows an interaction effect. The resulting
estimate is exponentiated so that it can be interpreted as the expected ratio
of costs resulting from a standard deviation change in the continuous
variable.


```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}
# ____ Marginal effects plots ----
# Custom wrapper of margins::margins for use with map_df
margins_custom <-
  function(mod, terms) { # Takes a model and a character vector of terms
    margins(mod, variables = terms, change = "sd") %>% summary %>% clean_names # Returns marginal effect of a 1 s.d. change in variable
  }

map_df(
  mods, 
  margins_custom,
  terms =
    c(
      "bankfull_width",
      "slope",
      "slope_deg",
      "n_culverts",
      "hdens_cat",
      "emp_const",
      "emp_agforest"
    ),
  .id = "model"
  ) %>%
  mutate(
    factor = case_when(
      factor == "bankfull_width" ~ "Bankfull width",
      factor == "slope" ~ "Stream slope",
      factor == "slope_deg" ~ "Terrain slope",
      factor == "n_culverts" ~ "Number of culverts",
      factor == "hdens_cat" ~ "Housing density",
      factor == "emp_const" ~ "Construction employment",
      factor == "emp_agforest" ~ "Ag/forestry employment"
    ),
    group = case_when(
      factor %in% c("Bankfull width", "Stream slope") ~ "Stream features",
      factor %in% c("Terrain slope", "Number of culverts") ~ "Others",
      factor %in% c("Construction employment", "Ag/forestry employment", "Housing density") ~ "Population features",
    ),
    estimate = exp(ame),
    conf.low = exp(lower),
    conf.high = exp(upper),
    p.value = I(p < 0.05),
    mod.color = case_when(
      p.value == TRUE & model == "mod_full" ~ "sig-pref",
      p.value == TRUE & model != "mod_full" ~ "sig-nopref",
      p.value == FALSE & model == "mod_full" ~ "nosig-pref",
      p.value == FALSE & model != "mod_full" ~ "nosig-nopref"
    ),
    mod.color = ordered(mod.color, levels = c("sig-pref", "sig-nopref", "nosig-pref", "nosig-nopref"))
  ) %>%
  select(model, factor, estimate, conf.low, conf.high, mod.color, group) %>%
  ggplot() +
  geom_pointrange(
    aes(
      y = reorder(factor, estimate),
      x = estimate,
      xmin = conf.low,
      xmax = conf.high,
      group = model,
      color = mod.color
    ),
    position = position_dodge(width = 0.8)
  ) +
  scale_color_manual(values = c("darkgreen", "darkolivegreen3", "grey60", "grey80")) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    y = NULL,
    x = "Project average costs", 
    title = "Marginal effects for continuous variables",
    subtitle = "Project average costs relative to a single standard deviation shift",
    caption = "Lines indicate 95% confidence interval; Preferred model highlighted in dark; Significant coefficients highlighted in color"
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot"
  ) +
  facet_wrap("group", ncol = 1, scales = "free_y")
```

A standard deviation increase in terrain slope and number of culverts 
are associated with lower average costs by about ten percent and half respectively in the
preferred model. The number of culverts result suggests evidence of returns
to scale. The cost-reducing effect of terrain slope is unituitive, and may be
picking up other cost-reducing effects in more remote, rugged, isolated
areas.  

The average marginal effects for all three population features are
insignificant in the preferred model. However, in some several of the
alternative models, ag/forestry employment and housing density each have
slightly positive effects, suggesting that each increase costs. The housing
density effect might be evidence of increased access costs due to negotiating
with multiple landowners. The employment effect is less intuitive, but may
also be picking up on cost increases in more rural areas, rather than a true
population effect.  

Bankfull width and stream slope both significant positive effects in the
preferred models. A standard deviation increase in either is associated with about
fifty percent higher average costs when the other is held at its mean. As seen
in the earlier figure and raw coefficients, this effect is driven largely by
the interaction term in the model. Note that the models for which the effects
are insignificant are those estimated using only Washington data, over either
the Washington basins only or WA RCO projects.  

## Fixed effects and categorical variables  
  
We can examine the fixed effect estimates to compare relative expected costs
across groups. We do this by exponentiating the point estimates. The
resulting value can be interpreted as the ratio of costs in a given group
relative to a base group


```{r }
# ____ Fixed effects plots ----
```

### Year effects  

```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}

# Year FE
map_df(mods[-4], tidy, conf.int = TRUE, .id = "model") %>%
  filter(str_detect(term, "year")) %>%
  mutate(
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high),
    year = str_remove(term, "factor[(]project_year[)]"),
    p.value = I(p.value < 0.05),
    mod.color = case_when(
      p.value == TRUE & model == "mod_full" ~ "sig-pref",
      p.value == TRUE & model != "mod_full" ~ "sig-nopref",
      p.value == FALSE & model == "mod_full" ~ "nosig-pref",
      p.value == FALSE & model != "mod_full" ~ "nosig-nopref"
    ),
    mod.color = ordered(mod.color, levels = c("sig-pref", "sig-nopref", "nosig-pref", "nosig-nopref"))
  ) %>%
  select(model, year, estimate, conf.low, conf.high, mod.color) %>%
  # left_join(mods_years_counts, by = c("year" = "completed_year")) %>%
  # mutate(
  #   year_n = paste0(
  #     year,
  #     "\n(n = ",
  #     comma(n),
  #     ")"
  #   )
  # ) %>%
  ggplot() +
  geom_pointrange(
    aes(
      x = year,
      y = estimate,
      ymin = conf.low,
      ymax = conf.high,
      group = model,
      color = mod.color
    ),
    position = position_dodge(width = 0.8)
  ) +
  scale_color_manual(values = c("darkgreen", "darkolivegreen3", "grey60", "grey80")) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = "Year", 
    y = "Project average costs", 
    title = "Year fixed effects",
    subtitle = "Project average costs relative to 2001 levels",
    caption = "Lines indicate 95% confidence interval; Preferred model highlighted in dark; Significant coefficients highlighted in color"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.75
    ),
    plot.title.position = "plot"
  )
```

For the preferred model, project average costs are nearly twice as high than
2001 levels between 2004 and 2010, with the exception of 2008, when other
factors are controlled for. Other models generally follow this pattern, while the
models for Western Washington basins/sources have fairly extreme standard
errors caused by small sample size.  

### Reporting source effects  

```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}

# Source FE
plotdata <- map_df(mods, tidy, conf.int = TRUE, .id = "model") %>%
  filter(str_detect(term, "source")) %>%
  mutate(
    source = str_remove(term, "project_source"),
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high),
    p.value = I(p.value < 0.05),
    mod.color = case_when(
      p.value == TRUE & model == "mod_full" ~ "sig-pref",
      p.value == TRUE & model != "mod_full" ~ "sig-nopref",
      p.value == FALSE & model == "mod_full" ~ "nosig-pref",
      p.value == FALSE & model != "mod_full" ~ "nosig-nopref"
    ),
    mod.color = ordered(mod.color, levels = c("sig-pref", "sig-nopref", "nosig-pref", "nosig-nopref"))
  ) %>%
  group_by(source) %>%
  select(model, source, estimate, conf.low, conf.high, mod.color)
ggplot() +
  geom_pointrange(
    aes(
      y = reorder(source, estimate),
      x = estimate,
      xmin = conf.low,
      xmax = conf.high,
      group = model,
      color = mod.color
    ),
    position = position_dodge(width = 0.6),
    data = plotdata
  ) +
  # geom_text(
  #   aes(
  #     x = c(0.5, 1.5),
  #     y = c(0.6, 0.6),
  #     hjust = c(1, 0),
  #     label = c("← Lower costs", "Higher costs →")
  #   ),
  #   size = 3.5
  # ) +
  scale_color_manual(values = c("darkgreen", "darkolivegreen3", "grey60", "grey80")) +
  # scale_x_continuous(breaks = waiver()) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    y = NULL, 
    x = "Project average costs", 
    title = "Reporting source fixed effects",
    subtitle = "Project average costs relative to OWRI",
    caption = "Lines indicate 95% confidence interval; Preferred model highlighted in dark; Significant coefficients highlighted in color"
  ) +
  # theme_clean() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    plot.background = element_rect(color = NA),
    plot.title.position = "plot",
    plot.caption.position = "plot"
  )
```

WDOT projects have average costs over five times as large as OWRI projects,
even when accounting for other factors. The standard errors on this estimate
are quite large though, suggesting a more accurate estimate might be possible
with more data from this source. Other sources lie BLM and REO have tightly
estimated positive effects. OWRI projects have among the lowest average costs.  

### Basin effects  

```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}

# Basin FE
map_df(mods, tidy, conf.int = TRUE, .id = "model") %>%
  filter(str_detect(term, "basin")) %>%
  mutate(
    basin = str_remove(term, "basin"),
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high),
    p.value = I(p.value < 0.05),
    mod.color = case_when(
      p.value == TRUE & model == "mod_full" ~ "sig-pref",
      p.value == TRUE & model != "mod_full" ~ "sig-nopref",
      p.value == FALSE & model == "mod_full" ~ "nosig-pref",
      p.value == FALSE & model != "mod_full" ~ "nosig-nopref"
    ),
    mod.color = ordered(mod.color, levels = c("sig-pref", "sig-nopref", "nosig-pref", "nosig-nopref"))
  ) %>%
  group_by(basin) %>%
  select(model, basin, estimate, conf.low, conf.high, mod.color) %>%
  ggplot() +
  geom_pointrange(
    aes(
      y = reorder(basin, estimate),
      x = estimate,
      xmin = conf.low,
      xmax = conf.high,
      group = model,
      color = mod.color
    ),
    position = position_dodge(width = 0.6)
  ) +
  scale_color_manual(values = c("darkgreen", "darkolivegreen3", "grey60", "grey80")) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    y = NULL, 
    x = "Project average costs", 
    title = "Basin fixed effects",
    subtitle = "Project average costs relative to Southern Oregon Coastal",
    caption = "Lines indicate 95% confidence interval; Preferred model highlighted in dark; Significant coefficients highlighted in color"
  ) +
  # theme_clean() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    plot.background = element_rect(color = NA),
    plot.title.position = "plot",
    plot.caption.position = "plot"
  )
```

After controlling for other factors with the preferred model, projects in the
Lower Snake Basin have the highest average costs, followed by the Middle
Columbia Basin, with average costs roughly four and three times those in the
Southern Oregon Coastal Basin. The Western Washington basins (Puget Sound and
Washington Coastal) both have typical average costs about double those in the
reference basin. These effects are tightly estimated, though the basins with
the smallest standard errors are Nothern Oregon Coastal and Willamette.  

### Land cover effects  

```{r }
# Land class
```
```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}

map_df(mods, tidy, conf.int = TRUE, .id = "model") %>%
  filter(str_detect(term, "nlcd")) %>%
  mutate(
    nlcd_current = str_sub(term, 27),
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high),
    p.value = I(p.value < 0.05),
    mod.color = case_when(
      p.value == TRUE & model == "mod_full" ~ "sig-pref",
      p.value == TRUE & model != "mod_full" ~ "sig-nopref",
      p.value == FALSE & model == "mod_full" ~ "nosig-pref",
      p.value == FALSE & model != "mod_full" ~ "nosig-nopref"
    ),
    mod.color = ordered(mod.color, levels = c("sig-pref", "sig-nopref", "nosig-pref", "nosig-nopref"))
  ) %>%
  group_by(nlcd_current) %>%
  select(model, nlcd_current, estimate, conf.low, conf.high, mod.color) %>%
  ggplot() +
  geom_pointrange(
    aes(
      y = reorder(nlcd_current, estimate),
      x = estimate,
      xmin = conf.low,
      xmax = conf.high,
      group = model,
      color = mod.color
    ),
    position = position_dodge(width = 0.6)
  ) +
  scale_color_manual(values = c("darkgreen", "darkolivegreen3", "grey60", "grey80")) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    y = NULL, 
    x = "Project average costs", 
    title = "NLCD land cover effects",
    subtitle = "Project average costs relative to Developed",
    caption = "Lines indicate 95% confidence interval; Preferred model highlighted in dark; Significant coefficients highlighted in color"
  ) +
  # theme_clean() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    plot.background = element_rect(color = NA),
    plot.title.position = "plot",
    plot.caption.position = "plot"
  )
```

Work sites in areas with the land cover classifications in the developed
group (the reference level) have among the highest costs, though those in the
planted-cultivated and water groups have high costs as well. Work sites in
forests and herbaceous land cover areas have average costs around a quarter
lower than the reference group.  

### Road feature effects  

```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}

# Road features
map_df(mods, tidy, conf.int = TRUE, .id = "model") %>%
  filter(str_detect(term, "here")) %>%
  mutate(
    here_var = case_when(
      term == "here_pavedY" ~ "Road paved (dummy)",
      str_detect(term, "factor") ~ str_replace(term, "factor[(]here_class[])]", "Road class: ")
      ),
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high),
    p.value = I(p.value < 0.05),
    mod.color = case_when(
      p.value == TRUE & model == "mod_full" ~ "sig-pref",
      p.value == TRUE & model != "mod_full" ~ "sig-nopref",
      p.value == FALSE & model == "mod_full" ~ "nosig-pref",
      p.value == FALSE & model != "mod_full" ~ "nosig-nopref"
    ),
    mod.color = ordered(mod.color, levels = c("sig-pref", "sig-nopref", "nosig-pref", "nosig-nopref"))
  ) %>%
  group_by(here_var) %>%
  select(model, here_var, estimate, conf.low, conf.high, mod.color) %>%
  ggplot() +
  geom_pointrange(
    aes(
      y = here_var,
      x = estimate,
      xmin = conf.low,
      xmax = conf.high,
      group = model,
      color = mod.color
    ),
    position = position_dodge(width = 0.6)
  ) +
  scale_color_manual(values = c("darkgreen", "darkolivegreen3", "grey60", "grey80")) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    y = NULL, 
    x = "Project average costs", 
    title = "Road feature effects",
    subtitle = "Project average costs relative to class 2",
    caption = "Lines indicate 95% confidence interval; Preferred model highlighted in dark; Significant coefficients highlighted in color"
  ) +
  # theme_clean() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    plot.background = element_rect(color = NA),
    plot.title.position = "plot",
    plot.caption.position = "plot"
  )
```

The only consistent effect found among the road feature variables is its
paved status. Paved roads are associated with work sites roughly 50 percent higher than unpaved work sites.  

### Scope and scale effects  

```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}

# Project scope effects
map_df(mods[-14], tidy, conf.int = TRUE, .id = "model") %>%
  filter(str_detect(term, "action")) %>%
  mutate(
    action_var = case_when(
      str_detect(term, "culvrem") ~ "Culvert removal (dummy)",
      str_detect(term, "culvinst") ~ "Culvert installation (dummy)"
    ),
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high),
    p.value = I(p.value < 0.05),
    mod.color = case_when(
      p.value == TRUE & model == "mod_full" ~ "sig-pref",
      p.value == TRUE & model != "mod_full" ~ "sig-nopref",
      p.value == FALSE & model == "mod_full" ~ "nosig-pref",
      p.value == FALSE & model != "mod_full" ~ "nosig-nopref"
    ),
    mod.color = ordered(mod.color, levels = c("sig-pref", "sig-nopref", "nosig-pref", "nosig-nopref"))
  ) %>%
  group_by(action_var) %>%
  select(model, action_var, estimate, conf.low, conf.high, mod.color) %>%
  ggplot() +
  geom_pointrange(
    aes(
      y = action_var,
      x = estimate,
      xmin = conf.low,
      xmax = conf.high,
      group = model,
      color = mod.color
    ),
    position = position_dodge(width = 0.6)
  ) +
  scale_color_manual(values = c("darkgreen", "darkolivegreen3", "grey60", "grey80")) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    y = NULL, 
    x = "Project average costs", 
    title = "Project scope effects",
    subtitle = "Project average costs relative to a project with only culvert improvements",
    caption = "Lines indicate 95% confidence interval; Preferred model highlighted in dark; Significant coefficients highlighted in color"
  ) +
  # theme_clean() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    plot.background = element_rect(color = NA),
    plot.title.position = "plot",
    plot.caption.position = "plot"
  )
```

There is some evidence from the models that installations are
more expensive than improvements, which are more expensive than removals,
though this effect largely washes out when the full suite of fixed effects is
included.  

# Benefit - cost visualizations  

```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}

# ____ Benefits and costs plots ----
```

As a simple examination of what kind of decision rule might be in play for
determining which culverts were selected, we plot as a benefit proxy total
upstream length (in km) versus observed cost per culvert. Under a
cost-targeting rule, all work sites would be to the left of a cost threshold,
while for a benefit-targeting rule all would be above a benefit threshold. A
benefit-cost ratio standard would be above an upward sloping line.  

```{r echo=F, message=F, warning=F, fig.dim=c(8,8)}
df_culv %>%
  mutate(
    # basin = fct_lump_lowfreq(basin, "OTHER"),
    basin = fct_lump_min(basin, 35, other_level = "OTHER")
  ) %>%
  ggplot() + 
  aes(
    x = cost_per_culvert,
    y = upst_dist,
    # y = tot_stream_length,
    color = project_year
  ) + 
  geom_point() +
  scale_y_log10("Total upstream length (km)", label = label_comma(1)) +
  scale_x_log10("Cost per culvert (K $USD)", label = label_dollar(1, scale = 0.001)) +
  scale_color_fermenter("Project year", palette = "Spectral", show.limits = TRUE, guide = guide_colorsteps(barwidth = 10, title.position = "top")) +
  theme(
    legend.position = "bottom",
    plot.title.position = "plot"
  ) +
  facet_wrap("basin") +
  ggtitle("Work sites in cost - benefit space", "Both on a log scale for clarity")
```

No evidence that projects are selected on cost or benefit-cost ratio basis.
It does somewhat appear that the observed projects follow a benefit targeting
pattern with a fairly low benefit cut-off, as evidenced by the horizontal
pattern among the data. For the Southern Oregon Coastal and Willamette
Basins, there appears to be a slight upward tilt in the higher-cost region,
indicating that benefit-cost targeting may be more frequent for higher-cost projects.  

Adding culverts where no project is observed with costs predicted via the
above models could reveal where observed projects exisit in the space
relative to the universe of potential projects, which should provide more
clarity to the above analysis.  

```{r echo=F, warning=F, message=F}
```

# Conclusions  
## Key findings  
  

```{r }
# Key findings ----
```

1. Stream features slope and bankfull width increase average costs, especially when they are both high.  
2. Paved roads are more expensive to improve, while other road variables have no discernible effect.  
3. Strangely, terrain slope at work site has a negative effect. Could this be picking up on some other 
factor associated with more rugged or remote work sites?  
4. Some evidence of economies of scale, in that work sites associated with
projects associated with more culverts tend to have lower average costs.  
  

```{r }
```

## Next steps  
  

```{r }
# Next steps ----
```

1. More variables: density of culvert project measures (issues w/
inconsistent reporting across space), distance to pop. center (euclidean or
routed with HERE roads from Blake), alternative "terrain ruggedness" measures, etc.  
2. Improved benefit estimates: total upstream distance by species and habitat use, weighted by catchment road density, etc.  
3. Forecast costs/benefits for culvert inventories from Oregon and Washington.  
4. Integrate Lorenz curve and Gini coefficient analysis from Babcock et al. (1997).  
5. Dig more into potential pitfalls of assigning project-level costs to work site-level observations.  
6. Incorporate Robby's findings on spatially dependent error structures.  


```{r end}
```

# End Matter  

```{r }
# Created on...
Sys.time()

# Using...
sessionInfo()

# Render output
# rmarkdown::render(here::here("R/C.culvertsspatial/07.spatialcostmodels.R"), output_file = here::here("output/culvertsmodels_report_2020oct06.html"))
```

